---
title: "GAE7"
author: "Your name here"
format: 
  html:
    theme: custom.scss
    embed-resources: true
    toc: true
    code-tools: true

csl: ecology.csl

editor: visual
bibliography: references.bib
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)

library(broom)
library(patchwork)
library(ggfortify)
library(latex2exp)
library(kableExtra)
library(tidyverse)

options(knitr.kable.NA = "")

```

## Scenario 1: Shakira the Saurian's "hips don't lie"

Establishing a dependable means to determine the sex of dinosaurs is one of the most sought after goals in dinosaur vertebrate paleontology [@Erickson2005]. Paleontologists would like to be able to determine the sex of dinosaurs from fossilized skeletal features. Living crocodilians belong to the Archosauria clade and are in the sister clade to the group including the Dinosauromorpha. Beginning in the 1950s, researchers have used living crocodilians as surrogates to look for sex-specific skeletal attributes that may prove useful in determining sex of fossil dinosaurs [@Romer1956].

### Methods

@Prieto-Marquez.etal2007 measured several characters that are thought to distinguish the sexes in American alligators, *Alligator mississipiensis*, one of the largest surviving reptiles. One of the characters measured was pelvic canal width (cm, *"pelvicwidth"*). The researchers also knew that it is common for anatomical features to be strongly correlated, so they also measured snout-vent length (m, *"snoutvent"*) to use as a covariate to correct for the size of alligator when comparing pelvic canal width among individuals.

#### Statistical Methods

```{r }
#| label: dataImportAlligators

alligator <- read_csv("alligator.csv", show_col_types = FALSE)
```

## Scenario 2:

### Introduction

@Russell.etal2017a reported on a 12-year experiment to examine how trees with different species traits may affect soil cation dynamics. One of the species traits they were interested in was fine root growth. Their rationale was, "Fine-root growth allows for exploration of the soil volume; thus differences among species in this trait could result in differential uptake of available soil cations."

### Methods

The authors reported that this experiment was begun in 1988 on abandoned pasture land that had once supported broad-leaved rain forest, now part of La Selva Biological Station, Costa Rica [@Russell.etal2017a]. The experiment consisted of five treatments: monoculture plantings of four native rain forest trees with differing life-history traits (see Table 1 in @Russell.etal2017a) and controls with no trees planted. Plots were 50 x 50 m (0.25 ha).

Treatments were deployed in a randomized complete block design with four spatial blocks. The authors report that the area is "hilly" and that each block was centered on a hilltop [@Russell.etal2017a]. The authors report that October of 2009, the *Vochysia guatemalensis* treatment in Block 1 was destroyed by a "stand-level lightening event" [@Russell.etal2017b], so there are only 19 observations total. See @Russell.etal2017a for details on fine root growth determination.

#### Statistical Methods

```{r }
#| label: dataImportRoots

# Species in monodominant plantations:  Hieronyma alchorneoides = HIAL;
# Pentaclethra macroloba = PEMA; Virola koschnyi = VIKO; and Vochysia
# guatemalensis = VOGU.  Control = unplanted, naturally regenerating treatment.

# Growth units: cm/cm2/yr

roots <- read_csv(
  "Russell_etal2017.csv",
  show_col_types = FALSE
  )
```

The authors provided their data via the Dryad Digital Data Repository [@Russell.etal2017b].

## Results

## References
